Studies of quantum dots

### 2017-01-23 Manuscript/paper.tex
Update spEnergiesQD/Manuscript/paper.tex
Add new figures

\input{head_plain.tex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{siunitx}
\usepackage{subfigure}

@@

Since for this paper we are interested in the addition and removal energies, which involve one particle more or one particle less than a closed-shell state, we can take the label $u$ to be simply the label of the state that is added or removed from the closed-shell state $\Phi$.  With this realization, we can then express the perturbation expansion in terms of summations over matrix elements as we had done before for the IM-SRG flow equation, perhaps with the aid of Wick's theorem or diagrammatic techniques.  This leads to the following expression for the second-order correction:
\begin{align*}
  \epsilon_p^{(2)}
  &=
  \frac{1}{2} \sum_{i a b} \frac{n_i \bar{n}_a \bar{n}_b |H^{[2]}_{p i a b}|^2}{H^{[1]}_p + H^{[1]}_i - H^{[1]}_a - H^{[1]}_b} \\
  &{}-
  \frac{1}{2} \sum_{i j a} \frac{n_i n_j \bar{n}_a |H^{[2]}_{i j p a}|^2}{H^{[1]}_i + H^{[1]}_j - H^{[1]}_p - H^{[1]}_a}
\end{align*}
[[write out the full thing using $H^1_{p,p}$]]
The expression is depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous terms in the third-order correction $\epsilon_p^{(3)}$, they are listed exclusively in a diagrammatic form in Figure \ref{fig:diagrams-sfe}.

@@

\begin{figure}
  \centering
  \subfigure[Results for $N=6$ and $\omega=1.0$]{
    \label{fig:N6hw1}
    \includegraphics[width=0.38\linewidth]{figures/6parthw1.pdf}
  }
  \subfigure[Results for $N=6$ and $\omega=0.5$]{
    \label{fig:N6hw05}
    \includegraphics[width=0.38\linewidth]{figures/6parthw05.pdf}
  } \\
  \subfigure[Results for $N=6$ and $\omega=0.28$]{
    \label{fig:N6hw028}
    \includegraphics[width=0.38\linewidth]{figures/6parthw028.pdf}
  }
  \subfigure[Results for $N=6$ and $\omega=0.1$]{
    \label{fig:N6hw01}
    \includegraphics[width=0.38\linewidth]{figures/6parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG) for circular quantum dots for $N=6$ and $\omega=1.0$ The results are compared with diffusion Monte Carlo (DMC), coupled cluster at the level of singles and doubles (CCSD), full configuration interaction (FCI) and Hartree-Fock calculations (HF) as functions of the number of major oscillator shells $R$. A harmonic oscilaltor basis has been used, with an unrenormalized Coulomb repulsion and the White generator. All energies in atomic units.}
  \label{fig:N6}
\end{figure}

\begin{table}
  \centering
  \caption{Ground state energy (in atomic units) of a circular quantum dot, with $N$ particles and oscillator frequency $\omega$. The IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$ results which are for $R=18$), employing a Hartree-Fock basis, a bare Coulomb interaction and the White generator. For $N=42$ and $\omega=0.1$ results are presented up to $R=22$ major shells. The results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained using the extrapolation formula of [[equation?]].}
  \label{tab:SRGDMC-R20}
  \begin{tabular}{SSSSS}
    \toprule
    {$N$} & {$\omega$} & {IM-SRG(2)} & {IM-SRG(2)$_{R\rightarrow \infty}$} & {DMC} \\
    \midrule
    6  & 1.0  &  20.1583  &  20.1468 &  20.1593(1) \\
       & 0.5  &  11.7690  &  11.7631 &  11.7848(1) \\
       & 0.28 &   7.5713  &   7.5695 &   7.6002(1) \\
       & 0.1  &   3.4962  &   3.4977 &   3.5539(1) \\
    \midrule
    12 & 1.0  &  65.7280  &  65.6990 &  65.7001(1) \\
       & 0.5  &  39.1631  &  39.1473 &  39.1596(1) \\
       & 0.28 &  25.6212  &  25.6140 &  25.6358(1) \\
       & 0.1  &  12.2223  &  12.2234 &  12.2698(1) \\
    \midrule
    20 & 1.0  & 155.9737  & 155.9130 & 155.8822(1) \\
       & 0.5  &  93.9219  &  93.8853 &  93.8752(1) \\
       & 0.28 &  61.94362 &  61.9240 &  61.9268(1) \\
       & 0.1  &  29.95263 &  29.9565 &  29.9779(1) \\
    \midrule
    30 & 1.0  & 308.7673  & 308.6053 & 308.5627(2) \\
       & 0.5  & 187.1671  & 186.8340 & 187.0426(2) \\
       & 0.28 & 124.0410  & 123.7171 & 123.9683(2) \\
       & 0.1  &  60.4319  &  60.1158 &  60.4205(2) \\
    \midrule
    42 & 1.0  & 543.3398  & 543.3304 & 542.9428(8) \\
       & 0.5  & 330.8885  & 330.8425 & 330.6306(2) \\
       & 0.28 & 220.0226  & 219.9534 & 219.8426(2) \\
       & 0.1  & 107.7851  & 106.8074 & 107.6389(2) \\
    \midrule
    56 & 1.0  & 880.4159  &          & 879.3986(6) \\
       & 0.5  & 538.8199  &          & 537.353(2)  \\
       & 0.28 & 360.83689 &          & 358.145(2)  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

[[Sarah's discussions on ground state energy are absent :(]]
In this section, we discuss the IM-SRG results for the ground-state energies of closed-shell systems.  We start with the Wegner generator in subsection [[??]].  Numerical instabilities in the integration process, as well as a rather slow convergence as function of the number of oscillator shells, favors the use of a Hartree-Fock basis and the Coulomb interaction.  Since higher correlations still lead to stiff equation systems, we change in subsection [[??]] to the White generator. In particular, we give an overview of the IM-SRG(2) ground state energies for closed-shell systems up to $N=42$ particles and compare with other \textit{ab initio} many-body methods.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/fig-methods}
  \caption{A high-level view of the ways in which many-body methods in this paper can be combined to calculate ground state, addition, and removal energies, shown as a directed acyclic graph.}
  \label{fig:methods}
\end{figure}

Figure \ref{fig:methods} shows high-level view of the ways in which many-body methods in this paper can be combined to to compute the energy observables.  There are many possible paths through which one can walk in this directed graph to obtain the output observables.  For example, one can do HF to get an initial rough estimate, followed by IM-SRG to refine the ground state, and then EOM to refine the addition and removal energies.  This is in fact one of the approaches benchmarked in this paper.

>>>>

\input{head_plain.tex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{siunitx}
\usepackage{subcaption}

@@

Since for this paper we are interested in the addition and removal energies, which involve one particle more or one particle less than a closed-shell state, we can take the label $u$ to be simply the label of the state that is added or removed from the closed-shell state $\Phi$.  With this realization, we can then express the perturbation expansion in terms of summations over matrix elements as we had done before for the IM-SRG flow equation, perhaps with the aid of Wick's theorem or diagrammatic techniques.  This leads to the following expression for the second-order correction:
\begin{align*}
  \epsilon_p^{(2)}
  &=
  \frac{1}{2} \sum_{i a b} \frac{n_i \bar{n}_a \bar{n}_b |H^{[2]}_{p i a b}|^2}{H^{[1]}_p + H^{[1]}_i - H^{[1]}_a - H^{[1]}_b} \\
  &{}-
  \frac{1}{2} \sum_{i j a} \frac{n_i n_j \bar{n}_a |H^{[2]}_{i j p a}|^2}{H^{[1]}_{i, i} + H^{[1]}_{j, j} - H^{[1]}_{p, p} - H^{[1]}_{a, a}}
\end{align*}
The expression is depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous terms in the third-order correction $\epsilon_p^{(3)}$, they are listed exclusively in a diagrammatic form in Figure \ref{fig:diagrams-sfe}.

@@

\subsection{Methodology}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/fig-methods}
  \caption{A schematic view of the various ways in which many-body methods in this paper could be combined to calculate ground state, addition, and removal energies.}
  \label{fig:methods}
\end{figure}

Before we discuss the results, let us briefly outline the concrete methodology used by our numerical experiments.  There is a some degree of flexibility in which many-body methods can be applied, which is not evident from reading the description of each method alone.

The typical approach is to apply them in the order shown in Figure \ref{fig:methods}.  Such an ordering maximizes the benefits of each method: Hartree-Fock acts as an initial, crude procedure to \textit{soften} the Hamiltonian, followed by IM-SRG or CC to refine the ground state energy, and then finally QDPT or EOM to refine the addition and removal energies.

The arrows in the diagram designate the flow of numerical data, which are generally matrix elements of some operator.  For HF and IM-SRG, these are simply (evolved) Hamiltonian matrix elements.  For CC, they would be elements of the cluster operator $\hat{T}$.

There are several paths through which one can traverse in this diagram to obtain the output observables.  Among these possibilities, there exist (a) HF + IM-SRG + QDPT, (b) HF + IM-SRG + EOM, and (c) HF + CC + EOM.  All three paths are explored in this paper.  There are, of course, other paths, but we will not consider every single possibility here.

Moreover, there is also the possibility of omitting one or more of these steps in the process.  For example, one can omit HF, but continue with the remaining two steps.  While this is possible, our experience indicate that the HF plays a strong role in increasing the rate of convergence of the post-HF methods, and thereby increasing the accuracy and efficiency as well.  Hence, there is very little reason to omit HF.

All results in this paper shall discuss results in which HF is included as the first step of the calculation, and thus we will not explicitly state ``HF'' unless there is no post-HF method used, in which case we write ``HF only''.

Unlike ground state energy, whose systems are characterized only by the number of particles $N$ and the oscillator frequency $\omega$, the addition (or removal) energy depend on an additional quantity: the total orbital angular momentum $M_\ell$, defined as the sum of the $m_\ell$ of each particle.  This is because there can be multiple states with $N + 1$ particles with near-degenerate energy.  For our numerical results, we will study exclusively the addition energies such that $|M_\ell|$ is as small as possible subject to the constraint that the particle added lies within the next shell.  This means the $N + 1$ states we are interested in are those with $|M_\ell| = K_{\mathrm{F}} \bmod 2$, while the $N - 1$ states are those with $|M_\ell| = 1 - K_{\mathrm{F}} \bmod 2$, where $K_{\mathrm{F}}$ is the number of occupied shells.

\subsection{Tables and figures}

\begin{table}
  \centering
  \caption{Ground state energy with $N$ particles and oscillator frequency $\omega$. The IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$ results which are for $R=18$), employing a Hartree-Fock basis, a bare Coulomb interaction and the White generator. For $N=42$ and $\omega=0.1$ results are presented up to $R=22$ major shells. The results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained using the extrapolation formula of [[equation? TODO: redo the extrapolations]].}
  \label{tab:SRGDMC-R20}
  \begin{tabular}{SSSSS}
    \toprule
    {$N$} & {$\omega$} & {IM-SRG(2)} & {IM-SRG(2)$_{R\rightarrow \infty}$} & {DMC} \\
    \midrule
    6  & 1.0  &  20.1583  &  20.1468 &  20.1593(1) \\
       & 0.5  &  11.7690  &  11.7631 &  11.7848(1) \\
       & 0.28 &   7.5713  &   7.5695 &   7.6002(1) \\
       & 0.1  &   3.4962  &   3.4977 &   3.5539(1) \\
    \midrule
    12 & 1.0  &  65.7280  &  65.6990 &  65.7001(1) \\
       & 0.5  &  39.1631  &  39.1473 &  39.1596(1) \\
       & 0.28 &  25.6212  &  25.6140 &  25.6358(1) \\
       & 0.1  &  12.2223  &  12.2234 &  12.2698(1) \\
    \midrule
    20 & 1.0  & 155.9737  & 155.9130 & 155.8822(1) \\
       & 0.5  &  93.9219  &  93.8853 &  93.8752(1) \\
       & 0.28 &  61.94362 &  61.9240 &  61.9268(1) \\
       & 0.1  &  29.95263 &  29.9565 &  29.9779(1) \\
    \midrule
    30 & 1.0  & 308.7673  & 308.6053 & 308.5627(2) \\
       & 0.5  & 187.1671  & 186.8340 & 187.0426(2) \\
       & 0.28 & 124.0410  & 123.7171 & 123.9683(2) \\
       & 0.1  &  60.4319  &  60.1158 &  60.4205(2) \\
    \midrule
    42 & 1.0  & 543.3398  & 543.3304 & 542.9428(8) \\
       & 0.5  & 330.8885  & 330.8425 & 330.6306(2) \\
       & 0.28 & 220.0226  & 219.9534 & 219.8426(2) \\
       & 0.1  & 107.7851  & 106.8074 & 107.6389(2) \\
    \midrule
    56 & 1.0  & 880.4159  &          & 879.3986(6) \\
       & 0.5  & 538.8199  &          & 537.353(2)  \\
       & 0.28 & 360.83689 &          & 358.145(2)  \\
    \bottomrule
  \end{tabular}
\end{table}

[[TODO: table for addrm]]

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-by-num-shells-6-2-ground-0-normal}
  \caption{[[...]]}
  \label{fig:by-num-shells-6-2-ground-0-normal}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-by-freq-10-6-ground-0-normal}
  \caption{The behavior of ground state energy as a function of the oscillator frequency $\omega$.  The energy is normalized with respect to the Hartree-Fock values to make plot easier to see.  Lower frequency leads to stronger correlations and thereby a more difficult problem. [[the fact that CCSD and MP2 are really close may be just a coincidence, since they cross around 12 shells but then diverge again]]}
  \label{fig:by-freq-10-6-gs-0-normal}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}{0.5\linewidth}
    \centering
    \includegraphics[width=\linewidth]{fig-by-freq-10-6-add-0-normal}
    \caption{Addition energy}
    \label{fig:by-freq-10-6-add-0-normal}
  \end{subfigure}%
  \begin{subfigure}{0.5\linewidth}
    \centering
    \includegraphics[width=\linewidth]{fig-by-freq-10-6-rm-1-normal}
    \caption{Removal energy}
    \label{fig:by-freq-10-6-rm-1-normal}
  \end{subfigure}
  \caption{The behavior of addition and removal energy as a function of the oscillator frequency $\omega$.  Like Figure \ref{fig:by-freq-10-6-gs-0-normal}, the energy is normalized with respect to the Hartree-Fock values.}
\end{figure}

In this section, we discuss the numerical results for the ground state, addition, and removal energies of closed-shell systems for a selection of methods.

For the IM-SRG results

As foreshadowed, we find the Wegner generator described in subsection [[??]].  Numerical instabilities in the integration process, as well as a rather slow convergence as function of the number of oscillator shells, favors the use of a Hartree-Fock basis and the Coulomb interaction.  Since higher correlations still lead to stiff equation systems, we change in subsection [[??]] to the White generator. In particular, we give an overview of the IM-SRG(2) ground state energies for closed-shell systems up to $N=42$ particles and compare with other \textit{ab initio} many-body methods.

@@
